Make any stupid plot you want

### 2021-02-15 draft/sections/intro.tex
more specified about math:

 This work presents a mathematical description of how data is transformed into graphic representations. While many researchers have identified and described important aspects of visualization, they have specialized in such different ways as to not provide a model general enough to natively support the full range of data and visualization types many general purpose modern visualization tools may need to support. In this work, we present a framework for understanding visualization built on topology and group theory. Using this mathematical formalism, we can interpret and extend prior work and also develop new tools. We validate our model by using it to re-design artist and data access layer of Maptlotlib, a general purpose visualization tool.

>>>>

 This work presents a mathematical description of how data is transformed into graphic representations. While many researchers have identified and described important aspects of visualization, they have specialized in such different ways as to not provide a model general enough to natively support the full range of data and visualization types many general purpose modern visualization tools may need to support. In this work, we present a framework for understanding visualization as equivariant maps between topological spaces. Using this mathematical formalism, we can interpret and extend prior work and also develop new tools. We validate our model by using it to re-design artist and data access layer of Matplotlib, a general purpose visualization tool.

### 2021-02-15 draft/sections/math.tex
more specified about math:

%% whats the right model/abstraction for a visualization library? 
%% why? the role this math is serving - static type checking/ reflection
%% tracking/automatically inferring
%% bookend w/ how this math does the why 
%% if we deleted this, what would we lose? 
%% can express continuous functions & not just discrete points
\note{should this be in 3rd person passive?}\\
Visualization is generally thought of as structure preserving maps from data into graphics, and in this section we formally define that structure and how it is preserved via equivariant maps. We can then specify that a faithful visual mapping is structure preserving, and apply these constraints to visualizations we may want to develop or implement. We model the data, visual characteristic, and graphic stages of visualization, shown in figure~\ref{fig:artist_stages}, as topological structures that encapsulate types of variables and continuity; by doing so we can develop implementations that keep track of both in ways that let us distribute computation while still allowing assembly and dynamic update of the graphic.

>>>>

can express continuous functions & not just discrete points
Visualization is generally thought of as structure preserving maps from data into graphics, and in this section we formally define that structure and how it is preserved via equivariant maps. We can then specify that a faithful visual mapping is structure preserving, and apply these constraints to visualizations we may want to develop or implement. We model the data, visual characteristic, and graphic stages of visualization, shown in figure~\ref{fig:artist_stages}, as topological structures that encapsulate types of variables and continuity; by doing so we can develop implementations that keep track of both in ways that let us distribute computation while still allowing assembly and dynamic update of the graphic.

### 2021-02-15 draft/sections/next.tex
more specified about math:

The model supports simple composition of glyphs by overlaying glyphs at the same position, but more work is needed to define an operator where the fiber bundles have shared $\gbase_2 \hookrightarrow \gbase_1$ such that fibers could be pulled back over the subset. This complex operator is necessary for building semantically meaningful components such as interactive visualizations that update on shared \gbase, such as brush-linked views\cite{beckerBrushingScatterplots1987,bujaInteractiveData1991} and verifying constraints in multiple view systems \cite{hullmanKeeping2018}. The models simple addition supports axes as standalone artists with overlapping visual position encoding, but the complex operator would allow for keeping labels consistent with text when for example the horizontal and vertical positions of the data are changed. In summary, the proposed scope of work for the dissertation is
\begin{itemize}
    \item expansion of the mathematical framework to include complex addition
    \item mathematical formulation of a graphic with axes labeling
    \item implementation of data with 2D continuous topology and a compatible artist
    \item provisional mathematics and implementation of user level composite artists
    \item proof of concept domain specific user facing library 
\end{itemize}
Additionally, implementing the complex addition operator would allow for developing a mathematical formalism and prototype of how interactivity would work in this model. Other potential tasks for future work is implementing a data object for a non-trivial fiber bundle and exploiting the models section level formalism to build distributed data source models and concurrent artists. This could be pushed further to integrate with topological\cite{heineSurveyTopologybasedMethods2016} and functional \cite{ramsayFunctionalDataAnalysis2006a} data analysis methods. While this paper formulates the math in terms of topology and groups, the model lends itself to a categorical formulation\cite{fongInvitationAppliedCategory2019,milewskiCategoryTheoryProgrammers}.

>>>>

The model supports simple composition of glyphs by overlaying glyphs at the same position, but more work is needed to define an operator where the fiber bundles have shared $\gbase_2 \hookrightarrow \gbase_1$ such that fibers could be pulled back over the subset. This complex operator is necessary for building semantically meaningful components such as interactive visualizations that update on shared \gbase, such as brush-linked views\cite{beckerBrushingScatterplots1987,bujaInteractiveData1991} and verifying constraints in multiple view systems \cite{hullmanKeeping2018}. The models simple addition supports axes as standalone artists with overlapping visual position encoding, but the complex operator would allow for keeping labels consistent with text when for example the horizontal and vertical positions of the data are changed. In summary, the proposed scope of work for the dissertation is
\begin{itemize}
    \item expansion of the mathematical framework to include complex addition
    \item mathematical formulation of a graphic with axes labeling
    \item implementation of data with 2D continuous topology and a compatible artist
    \item provisional mathematics and implementation of user level composite artists
    \item proof of concept domain specific user facing library 
\end{itemize}
Additionally, implementing the complex addition operator would allow for developing a mathematical formalism and prototype of how interactivity would work in this model. Other potential tasks for future work is implementing a data object for a non-trivial fiber bundle and exploiting the models section level formalism to build distributed data source models and concurrent artists. This could be pushed further to integrate with topological\cite{heineSurveyTopologybasedMethods2016} and functional \cite{ramsayFunctionalDataAnalysis2006a} data analysis methods. While this paper formulates visualization in terms of monoidal action homomorphisms between fiberbundles, the model lends itself to a categorical formulation\cite{fongInvitationAppliedCategory2019,milewskiCategoryTheoryProgrammers} that could be further explored.
